Optimizing a PoS Tag Set for Dependency Parsing

### 2016-05-06 submission.tex
changes to paragraph on determiners in tag set experiments

\paragraph{Adjectives}
Adjectives agree with the noun they modify in terms of gender, number and
definiteness in Norwegian. Adjectives are also inflected for degree (positive,
comparative or superlative).

In Table \ref{adjresults}, we report the results of modifying the \texttt{pron}
tag in NDT. All features except for number led to increases in parser accuracy
scores, the most successful of which is degree with a reported LAS of 87.29\%,
while distinguishing adjectives on definiteness yield an LAS of 87.14\% and
gender leads to LAS of 87.10\%.

@@

Turning to combinations of features, definiteness and number achieve the best
results, very close to that of degree, with 0.02 percentage points lower LAS
and identical UAS. Adjectives agree with their head noun and determiner in
definiteness and number, making this an expected improvement. The combination
of definiteness and degree is also quite promising, obtaining LAS of 87.23\%
and UAS of 90.39\%. It is interesting that none of the combinations surpass the
experiment with degree alone, which indicates that degree does not interact
with the other features in any syntactically significant way.

@@

\paragraph{Determiners}
Like adjectives, determiners in Norwegian agree with the noun they modify in
terms of gender, number and definiteness. Possessive pronouns are
treated/analyzed as determiners in NDT.

%For a number of reasons, we do not report results from combinations of features
%for determiners, as the features could not be combined in any meaningful way.
%For instance, determiners in plural never have marked definiteness, hence this
%combination would only apply to determiners in singular and roughly correspond
%to the distinction on definiteness alone. Moreover, determiners in plural are
%not distinguished in terms of gender (thus ruling out the combination of number
%and gender), neither are definite determiners. Another aspect taken into
%consideration is that the most promising distinction, definiteness, applies to
%such a small number of tokens that more fine-grained distinctions would be
%overly sparse.

The results from the experiments with determiners are shown in Table
\ref{detresults}.  Introducing the distinction of type led to an increase in
tagger accuracy by 0.14 percentage points to 97.61\%, while marginally
impacting the parsing, with LAS of 87.00\%, 0.01 percentage points below that
of the original tag set.  The increase in tagger accuracy when introducing the
distinction of type is noteworthy, as we expected the finer granularity to lead
to a decrease in accuracy. This serves to indicate that more fine-grained
distinctions for determiners, which is a quite disparate category in the
treebank, may be quite useful for tagging. However, as it has negative impact
on parsing, we can conclude that the type of a determiner does not assist in
generalizing syntactic patterns. Most determiners, regardless of type, appear
in the same syntactic constructions (i.e., before a noun or noun phrase).

Gender, on the other hand, improved the parsing, but complicated the tagging,
as the various genders are often difficult to differentiate, especially so in
the case of masculine and feminine, which share many of the same determiners.
The number of a determiner, i.e., singular or plural, led to a small increase
in tagger accuracy and LAS, while marginally lower UAS, 90.18\%, 0.01
percentage points lower than that of the original tag set. The
introduction of definiteness to the determiners led to the best parsing
results, LAS of 87.30\% and UAS of 90.42\%, while also increasing the tagger
accuracy slightly. The increase in LAS and UAS is rather interesting, as there
are only 121 determiner tokens with marked definiteness in the development
data. As this accounts for a very small number of tokens, we did not consider
further fine-grained modifications with definiteness. This goes to show that
tokens with overt definiteness have noticeable impact on the syntactic parsing,
and that distinguishing on definiteness is very beneficial.

@@

\paragraph{Pronouns}
Pronouns in Norwegian include personal, reciprocal, reflexive and
interrogative. They can exhibit gender, number and person. Personal pronouns
have case (accusative or nominative).

>>>>

\paragraph{Adjectives}
Adjectives agree with the noun they modify in terms of gender, number and
definiteness in Norwegian. Furthermore, adjectives are inflected for
degree (positive, comparative or superlative).

In Table \ref{adjresults}, we report the results of modifying the \texttt{pron}
tag in NDT. All features except for number lead to increases in parser accuracy
scores, the most successful of which is degree with a reported LAS of 87.29\%,
while distinguishing adjectives on definiteness yields an LAS of 87.14\% and
introducing the distinction of gender leads to LAS of 87.10\%.

@@

Turning to combinations of features, definiteness and number achieve the best
parser accuracy scores, very close to those of degree, with 0.02 percentage
points lower LAS and identical UAS. Adjectives agree with their head noun and
determiner in definiteness and number, making this an expected improvement. The
combination of definiteness and degree is also quite promising, obtaining LAS
of 87.23\% and UAS of 90.39\%. It is interesting that none of the combinations
surpass the experiment with degree alone, which indicates that degree does not
interact with the other features in any syntactically significant way.

@@

\paragraph{Determiners}
Like adjectives, determiners in Norwegian agree with the noun they modify in
terms of gender, number and definiteness. We do not report results from
combinations of features for determiners, as the features could not be combined
in any meaningful way.
%VPossessive pronouns are
%treated/analyzed as determiners in NDT.

%For instance, determiners in plural never have marked definiteness, hence this
%combination would only apply to determiners in singular and roughly correspond
%to the distinction on definiteness alone. Moreover, determiners in plural are
%not distinguished in terms of gender (thus ruling out the combination of number
%and gender), neither are definite determiners. Another aspect taken into
%consideration is that the most promising distinction, definiteness, applies to
%such a small number of tokens that more fine-grained distinctions would be
%overly sparse.

The results from the experiments with determiners are shown in Table
\ref{detresults}. Introducing the distinction of type (demonstrative,
amplifier, quantifier, possessive or interrogative) led to an increase in
tagger accuracy of 0.14 percentage points to 97.61\%, while marginally
impacting the parsing, with LAS of 87.00\%, 0.01 percentage points below that
of the original tag set.  The increase in tagger accuracy when introducing the
distinction of type is noteworthy, as we expected the finer granularity to lead
to a decrease in accuracy. This serves to indicate that more fine-grained
distinctions for determiners, which is a quite disparate category in the
treebank, may be quite useful for tagging.

%However, as it has negative impact
%on parsing, we can conclude that the type of a determiner does not assist in
%generalizing syntactic patterns. Most determiners, regardless of type, appear
%in the same syntactic constructions (i.e., before a noun or noun phrase).

Gender, on the other hand, improved the parsing (87.09\% LAS), but complicated
the tagging, as the various genders are often difficult to differentiate,
especially so in the case of masculine and feminine, which share many of the
same determiners. The number of a determiner, i.e., singular or plural, led to
a small increase in tagger accuracy and LAS, while marginally lower UAS of
90.18\%, 0.01 percentage points lower than that of the original tag set. The
introduction of definiteness to the determiners led to the best parsing
results, LAS of 87.30\% and UAS of 90.42\%, while also increasing the tagger
accuracy slightly. The increase in LAS and UAS is rather interesting, as there
are only 121 determiner tokens with marked definiteness in the development
data. As this accounts for a very small number of tokens, we did not consider
further fine-grained modifications with definiteness. This serves to indicate
that tokens with overt definiteness have noticeable impact on the syntactic
parsing and that distinguishing on definiteness is very useful.

@@

\paragraph{Pronouns}
Pronouns in Norwegian are personal, reciprocal, reflexive and interrogative.
They can furthermore exhibit gender, number and person, while personal pronouns
can be distinguished case (accusative or nominative).
